---
title: "Cross Validation - Exercises"
author: "A very cool R Lady"
date: "13 October 2018"
output: html_document
editor_options: 
  chunk_output_type: console
---

# Setting up (Packages/Data/Etc)

```{r Packages}
library(class)
library(randomForest)
library(cvTools)
library(ggplot2)
library(caret)
```


```{r Setting up the Data}
data <- read.csv("data/diabetes.csv")
data$class <- as.factor(data$class) #remember, classification method require our response variable to be a factor!
```

```{r Explore the data}
head(data)
str(data)


# Seperate Responses
X <- data[,-9]
y <- data$class
```

```{r Setting up CV parameters}
n <- nrow(X)
K <- 5 # Choose 5 fold CV
```


# Exercise 1) CV loop for different values of k for KNN

```{r}
k.vals <- 1:50
# Insert code here
```


We can plot the error as a function of *k*.

```{r plotting}
# Create a data frame with k.vals and CV errors - then plot (using either base/ggplot)
```

# Exercise 2) CV loop for randomForest

```{r}

# Create a CV loop similar to the one shown in the slides for KNN, but for randomForest instead. Revise last weeks slides for a reminder on randomForest syntax.
```

We can see that randomForest is outperforming KNN.

# Exercise 3) Repeated CV loop for KNN (10 times) using the best K value

```{r repeated KNN CV loop}

reps <- 10

# Insert R code here
```



# Exercise 4) Repeated CV loop for randomForest

```{r repeated randomForest CV loop}

reps <- 10

# Insert R code here

```

Compare results between RF and KNN using a comparative boxplot

```{r quick compare}
# Hint, can do this very quickly using baseR!
```




# Exercise 5) Using the `caret` package and comparing results

```{r caret for KNN}
fitControl <- trainControl(## 5-fold CV
                           method = "repeatedcv",
                           number =5,
                           ## repeated ten times
                           repeats = 10)

set.seed(1)
knnFit1 <- train(class ~ ., data = data, 
                 method = "knn", 
                 trControl = fitControl)
knnFit1

cv.err.knn <- 1-knnFit1$results$Accuracy[3] #CV error is 1-Accuracy. We pick the 3rd accuracy as the 9 was picked as best parameter.
```

```{r caret for RF}

# Repeat above for RF.

```

```{r caret for GLM}

# Repeat above for GLM. Be careful, you will need an extra argument besides method!
```

```{r caret for rpart}

# Repeat above for rpart.
```